---
title: "Visualize"
author: "Vahab"
date: '2022-03-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 3.1 Read Data
```{r}
library(here) #Get paths using here
library(readr)
library(DT)

# variables
url_ac <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
# if ERDDAP server down (Error in download.file) with URL above, use this:
#    url_ac <- "https://raw.githubusercontent.com/noaa-iea/r3-train/master/data/cciea_AC.csv"
csv_ac <- here("data/cciea_AC.csv")

#download data if needed
if(!file.exists(csv_ac))
  download.file(url_ac, csv_ac)

#read data into a tibble

d_ac<- read_csv(csv_ac, col_names = F, skip = 2)
names(d_ac) <-  names(read_csv(csv_ac))

# show data
datatable(d_ac)

# here::here() function starts the path based on looking for the *.Rproj file in the current working directory or higher level folder. In this case it should be the same folder as your current working directory so seems unnecessary, but it’s good practice for other situations in which you start running Rmarkdown files stored in subfolders (in which case the evaluating R Chunks assume the working directory of the .Rmd).

# When using readr::read_csv()  columns of character type are not converted to type factor by default. It will also default to being read in as a tibble rather than just a data.frame

```

```{r}
library(dplyr)
library(ggplot2)

# subset data
d_coast <- d_ac %>% 
  # select columns
  select(time, total_fisheries_revenue_coastwide) %>% 
  # filter rows
  filter(!is.na(total_fisheries_revenue_coastwide))
datatable(d_coast)
```
### 3.2 Plot statically with ggplot2
#### 3.2.1 Simple line plot + geom_line()
```{r}
# ggplot object
p_coast <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line()

# show plot
p_coast
```
#### 3.2.2 Trend line + geom_smooth()
```{r}
p_coast + 
  geom_smooth(method = "lm")
```
#### 3.2.3 Distribution of values + geom_histogram()
```{r}
d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_histogram(bins = 10)
```
#### 3.2.4 Series line plot aes(color = region)
```{r}
# Next, let’s also show the other regional values (CA, OR and WA; not coastwide) in the plot as a series with different colors. To do this, we’ll want to tidy the data into long format so we can have a column for total_fisheries_revenue and another region column to supply as the group and color aesthetics based on aesthetics we see are available for geom_line():
library(stringr)
library(tidyr)

d_rgn <- d_ac %>%
  select(time, starts_with("total_fisheries_revenue")) %>%
  #exclude colulmn
  select(-total_fisheries_revenue_coastwide) %>%
  pivot_longer(-time) %>%
  #mutate region by stripping other words
  #str_replace::stringr replace matched patterns in a string
  mutate(
    region = name %>% 
      str_replace("total_fisheries_revenue_", "") %>%
      str_to_upper()) %>%
  # filter for not NA
  filter(!is.na(value)) %>% 
  # select columns
  select(time, region, value)

p_rgn <- ggplot(d_rgn, aes(x=time, y=value, group=region, color=region)) + geom_line()

p_rgn

```
#### 3.2.5 Update labels + labs()
```{r}
p_rgn <- p_rgn +
  labs(
    title  = "Fisheries Revenue",
    x      = "Year",
    y      = "Millions $ (year 2015)",
    color  = "Region")
p_rgn
```
#### 3.2.6 Multiple plots with facet_wrap()
##### When you want to look at similar data one variable at a time, you can use facet_wrap() to display based on this variable.
```{r}
p_rgn +
  facet_wrap(~region)
```
### 3.2.7 Bar plot + geom_col()
#### Another common visualization is a bar plot. How many variables does geom_bar() use versus geom_col()?
```{r}
library(glue)
library(lubridate)

yr_max <- year(max(d_rgn$time))

d_rgn %>% 
  # filter by most recent time
  filter(year(time) == yr_max) %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_col() +
  # add labels
  labs(
    title  = glue("Fisheries Revenue for {yr_max}"),
    x      = "Region",
    y      = "Millions $ (year 2015)",
    fill   = "Region")
```
### 3.2.8 Variation of series with + geom_boxplot()
```{r}
d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_boxplot() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
```
### 3.2.9 Variation of series with + geom_violin()
```{r}
p_rgn_violin <- d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_violin() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
p_rgn_violin
```
### 3.2.10 Change Theme theme()
#### We’ve already manipulated the theme() in dropping the legend. You can create your own theme or use some of the existing.

```{r}
p_rgn_violin +
  theme_classic()

```
### 3.3 Plot interactively with plotly or dygraphs
#### 3.3.1 Make ggplot interactive with plotly::ggplotly()
```{r}
#When rendering to HTML, you can render most ggplot objects interactively with plotly::ggplotly(). The plotly library is an R htmlwidget providing simple R functions to render interactive JavaScript visualizations.

plotly::ggplotly(p_rgn)
```
### 3.3.2 Create interactive time series with dygraphs::dygraph()
```{r}
# Another htmlwidget plotting library written more specifically for time series data is dygraphs. Unlike the ggplot2 data input, a series is expected in wide (not tidy long) format. So we use tidyr’s pivot_wider() first.

library(dygraphs)
d_rgn_wide <- d_rgn %>% 
  mutate(
    Year = year(time)) %>% 
  select(Year, region, value) %>% 
  pivot_wider(
    names_from  = region,
    values_from = value)
datatable(d_rgn_wide)

d_rgn_wide %>% 
  dygraph() %>% 
  dyRangeSelector()
```
### Further Reading
Introductory ggplot2 topics not yet covered above are:

Other plot types: scatter, area, polar, ….

Changing scales of axes, color, shape and size with scale_*() functions.

Transforming coordinate system, eg coord_flip() to swap x and y axes for different orientation.

Adding text annotations.

Changing margins.

Summarization methods with stat_*() functions.

Here are further resources:

Learning ggplot2 | ggplot2
ggplot2: Elegant Graphics for Data Analysis: online book by Hadley Wickham
3. Data visualisation | R for Data Science: chapter from online book by Hadley Wickham and Garrett Grolemund
R Graphics Cookbook, 2nd edition
